Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints

### 2017-09-25 2017/ParOptBook/Chapter4.tex
parOptBook: update chapter 4

As before, the objective function $\varphi$ (hereinafter denoted also as $g_{m+1}$) and the left parts of the constraints $g_j(x)\le 0,1\le j\le m$ re assumed to satisfy \emph{Lipschitz condition} with the corresponding constants $L_j,1\le j\le m+1$, namely:
\begin{equation}
  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2\in [a,b]
\end{equation}

@@

\emph{Rule 9.} Execute the next trial at the middle point of the interval $(x_{t-1},x_t )$ if the indices of its boundary points are not the same, i.e.,
\begin{equation}
  x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
\end{equation}
Otherwise, accept
\begin{equation}
\label{eq4:next_point_par}
x^{k+p} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu_\nu},\; \nu=\nu(x_{t-1})=\nu(x_t).
\end{equation}
as the point of the next trial.

@@

\begin{proof}
  Since $\overline x$ is a limit point, there exists a sequence of the trial numbers $q^1,q^2,\dots$ satisfying the conditions
  \begin{equation}
    x^{q+1},\overline x\in[x_{t-1},x_t], t=t(q),q\in\{q^p\},p=1,2,\dots.
  \end{equation}
  These conditions reflect the fact that the point of the $(q+1)^{th}$ trial falls into the interval $[x_{t-1},x_t]$ containing the point $\overline x$ at the step $q$, i.e., $t=t(q)$.
\end{proof}

>>>>

As before, the objective function $\varphi$ (hereinafter denoted also as $g_{m+1}$) and the left parts of the constraints $g_j(x)\le 0,1\le j\le m$ re assumed to satisfy \emph{Lipschitz condition} with the corresponding constants $L_j,1\le j\le m+1$, namely:
\begin{equation}
  \label{eq4:lip_condition}
  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2\in [a,b]
\end{equation}

@@

\emph{Rule 9.} Execute the next trial at the middle point of the interval $(x_{t-1},x_t )$ if the indices of its boundary points are not the same, i.e.,
\begin{equation}
  \label{eq4:next_point_1_par}
  x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
\end{equation}
Otherwise, accept
\begin{equation}
\label{eq4:next_point_2_par}
x^{k+p} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu_\nu},\; \nu=\nu(x_{t-1})=\nu(x_t).
\end{equation}
as the point of the next trial.

@@

\begin{proof}
  Since $\overline x$ is a limit point, there exists a sequence of the trial numbers $q^1,q^2,\dots$ satisfying the conditions
  \begin{equation}
    x^{q+1},\overline x\in[x_{t-1},x_t], t=t(q),q\in\{q^p\},p=1,2,\dots.
  \end{equation}
  These conditions reflect the fact that the point of the $(q+1)^{th}$ trial falls into the interval $[x_{t-1},x_t]$ containing the point $\overline x$ at the step $q$, i.e., $t=t(q)$.

  It follows from this condition, from the rules of the next trial point selection \eqref{eq4:next_point_1_par} and \eqref{eq4:next_point_2_par}, and from the estimate of the value $\mu_\nu$ from \eqref{eq4:lip_est_par}, that
  \begin{equation}
    \label{eq4:4_60}
    \max\{x_t-x^{q+1},x^{q+1}-x_{t-1}\}\le\gamma \Delta_t,
  \end{equation}
  where $\gamma=(r+1)/2r<1$ and
  \begin{equation}
    r=\min\{r_\nu:1\le\nu\le m+1\}>1.
  \end{equation}
  Then, the property (4.56) is a consequence of inequality \eqref{eq4:4_60}.

  At every step $k\ge 1$ there exists an interval $(x_{i-1},x_i), \:i=i(k)$, at one of the ends of which the value $z^*_\omega$ from \eqref{eq4:z_const_par} is achieved. This value corresponds to the index $\omega$ defined (when $\omega<m$) by the conditions $I_\omega\not=\O,\: I_{\omega+1}\not=\O$. If $I_{m+1}=\O$, the index $\omega=m+1$. If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according (4.47) and (4.50), $R(i)=2\Delta_i>0$. In the case when $\nu(x_{i -1} )= \nu(x_i )=\omega$, the estimate
  \begin{displaymath}
    \max(z_{i-1}, z_i)-z^*_\omega\le\mu\Delta_i
  \end{displaymath}
  follows from \eqref{eq4:lip_est_par} and, along with \eqref{eq4:characteristic_par}, gives the estimate
  \begin{displaymath}
    R(i)=\Delta_i\left(1-\frac{\max(z_{i-1},z_i)-z^*_\omega}{r_\omega\mu_\omega\Delta_i}\right)\ge (1-r^{-1})\Delta_i,
  \end{displaymath}
  where $r$ is from (4.61). Thus, at any search step $k\ge 1$ there exists an interval having a strictly positive characteristic. Hence, taking into account the rules of selection of the interval with the greatest characteristic \eqref{eq4:MaxR_par}, one can conclude that the intervals from system (4.54) should also have the positive characteristics since these ones undergo splitting up by the trial points falling into these intervals. Consequently, conditions (4.57) hold.

  It follows from the trial execution rules and condition (4.35) that for $\nu(x_i )\le m$ the values $z_i$ assigned to the points $x_i$ from \eqref{eq4:par_search_data} are positive. Therefore, according to \eqref{eq4:z_const_par},
  \begin{displaymath}
    z_i=g_\nu(x_i)\ge z^*_nu,\:1\le\nu=\nu(x_i)\le m+1,\: 0\le i\le k.
  \end{displaymath}
    Now from the rules \eqref{eq4:characteristic_par} for the characteristics and the limit (4.56), taking into account the condition \eqref{eq4:lip_condition} of Lipschitzness of the functions and inequality (4.57), one can derive out the truth of statement (4.58).
\end{proof}
\begin{lemma}
  (on the existence of a feasible point). Assume the conditions of Theorem \ref{th4:convergence_par} to be fulfilled. Then, there exists the trial index $h\ge 1$, for which
  \begin{equation}
    \label{eq4:62}
    \nu(x^h)=m+1.
  \end{equation}
\end{lemma}
\begin{proof}
  Let us assume the opposite and consider the interval $[x_{t-1},x_t]$ including the problem solution $x^*$ from \eqref{eq4:problem} at the step $k$, i.e., $t=t(k)$. Since we have assumed the equality \eqref{eq4:62} not to take place and, consequently, the sequence $\{x^k\}$ does not include the points from the feasible domain $Q$, the interval $[\alpha,\beta]$ from (4.52) should satisfy the conditions
  \begin{equation}
    x^*\in [\alpha,\beta]\subset[x_{t-1}, x_t].
  \end{equation}
  Then, from this condition as well as from the Lipschitz properties of the problem functions the relations
  \begin{gather}
    z_{t-1}\le L_\nu(\alpha - x_{t-1})\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t-1}) \\
    z_t\le L_\nu(x_{t}-\beta)\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t})
  \end{gather}
  follow. It was taken into account when deriving out of the above relations that
  \begin{displaymath}
    g_\nu(x)\le 0,\: x\in[\alpha,\beta]\subset Q.
  \end{displaymath}
  Analogously, the case $\nu=\nu(x_{t-1})=\nu(x_t)$ corresponds to the inequality
  \begin{equation}
    z_t+z_{t-1}\le L_\nu \Delta_t - L_\nu(\beta-\alpha).
  \end{equation}

  Now, from the characteristic rules \eqref{eq4:characteristic_par}, from inequalities (4.64)–(4.66), and from the inequality
  \begin{displaymath}
    z^*_\nu\ge-\epsilon_\nu
  \end{displaymath}
  resulting from \eqref{eq4:z_const_par} either the estimate
  \begin{displaymath}
    R(t)\ge\Delta_t(1-2L_\nu/r_\nu\mu_\nu)+2(L_\nu(\beta-\alpha)-2\epsilon_\nu)/r_\nu\mu_\nu,
  \end{displaymath}
  or the estimate
  \begin{displaymath}
    R(t)\ge\Delta_t(1-2L_\nu/r_\nu\mu_\nu)+4(L_\nu(\beta-\alpha)-2\epsilon_\nu)/r_\nu\mu_\nu,
  \end{displaymath}
  is true. Taking into account the third condition of Theorem \ref{th4:convergence_par} and the fourth one, in both cases we get the inequality
  \begin{equation}
    R(t(k))>0
  \end{equation}
  which holds for sufficiently large step numbers $k$.

  From (4.58), (4.67), and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from (4.63) follows. Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals. After that, at least one of those will include the point $x^*$. The repetition of this division generates (with $k\to\infty$) an infinite sequence of the subintervals shrinking to the point $x^*$ . In other words, if the third condition of Theorem \ref{th4:convergence_par} and the fourth one are satisfied, $x^*$ is the limit point of the sequence ${x^k}$.

  Since the coincidence $x^*$ with one of the trial points means (4.62) to hold, then it remains to consider the case when $x^*\in(x_{t- 1} ,x_t ),\: t=t(k)$ for any $k\ge 1$. Here, because of the first condition of Theorem \ref{th4:convergence_par} and the definition (4.31) either $[x_{t-1},x^*]\in Q$ or $[x^* ,x_t]\in Q$ if the number $k$ is sufficiently large. In this case, either $\nu(x_{t-1})=m+1$ or $\nu (x_t )=m+1.$
\end{proof}

\begin{lemma}
  on the feasibility of any limit point). Assume the conditions of Theorem \ref{th4:convergence_par} to be met. Then any limit point $\overline x$ of the sequence $\{x_k\}$ is feasible, i.e.,
  \begin{equation}
    \nu(\overline x)=m+1
  \end{equation}
\end{lemma}

\begin{proof}
  Let us assume the opposite. According to (4.55) and (4.56),
\end{proof}
